---
title: "Figures for manuscript"
output: html_document
date: "`r Sys.Date()`"
---

```{r plot_settings, echo = FALSE, message = FALSE, warning = FALSE }

library(tidyverse)
library(patchwork)
library(RColorBrewer)
library(quickpalette)
library(shades)
library(cowplot)

custom_cols <- c("A_H2O2"= "#FF7433", "A" = "#FFAB85", "B_H2O2" = "#467986", "B" = "#79ACB9")

A <- expression("MnSO"[4])
AH <- expression(paste("MnSO"[4]~"+ H"[2]*"O"[2]))
B <- expression("EDTA")
BH <- expression(paste("EDTA + H"[2]*"O"[2]))


custom_names <- c("A_H2O2"= AH,
                  "A" = A,
                  "B_H2O2" = BH,
                  "B" = B)

```

## H<sub>2</sub>O<sub>2</sub> concentration

**Figure 1 - H<sub>2</sub>O<sub>2</sub> optimization** We evaluated how the concentration of H<sub>2</sub>O<sub>2</sub> affected the oxidation of MBTH/DMAB using two concentrations of MnP (0.005 U and 0.008 U, 10 µl in the 200 µl reaction).

We found that both too low and too high concentrations were sub-optimal for the reactions and that a lower concentration (2.5 mM) than we usually have been using (5 mM) seemed to be optimal for our assay conditions

```{r H2O2, echo = FALSE, message = FALSE, warning = FALSE}
h2o2.dat <- read_csv("Data/Clean/clean_H2O2_concentration.csv")

h2o2.dat$H2O2_conc <- as.factor(h2o2.dat$H2O2_conc)
h2o2.dat$MnP_conc <- as.factor(h2o2.dat$MnP_conc)

# Summarize mean values for technical replicates for each time point

h2o2.sum <- h2o2.dat |> 
  group_by(time_min, MnP_conc, H2O2_conc) |> 
  summarise(mean = mean(absorbance))

#remove half of the points to reduce clutter to the figure
h2o2.sum.reduced<-h2o2.sum[-grep(".5",h2o2.sum$time_min,fixed = TRUE),]

#rename from mM H2O2 conc to µM if desired
#h2o2.sum.reduced$H2O2_conc<-recode(h2o2.sum.reduced$H2O2_conc, "0" = "0",
#                                                                "0.375" = "18.75",
 #                                                               "0.75" = "37.50",
#                                                                "1.25" = "62.50",
#                                                                "2.5" = "125",
#                                                                  "5" = "250",
#                                                                  "10" = "500",
#                                                                  "50" = "2500")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

grad_cols <- c("0"= "#A1C4CE", "0.375" = "#87B3C0", "0.75" = "#6CA3B2", "1.25" = "#5490A0", "2.5" = "#467886", "5"= "#38606B", "10" = "#2A4850", "50" = "#1C3035" )

grad_cols2 <- c("0"= "#60ae94", "0.375" = "#6c9473", "0.75" = "#568678", "1.25" = "#506c78", "2.5" = "#436074", "5"= "#334151", "10" = "#1b283b", "50" = "#0a0c1a" )

grad_cols3 <- c("0"= "#1ee19b", "0.375" = "#2cbd89", "0.75" = "#0f9774", "1.25" = "#009657", "2.5" = "#226346", "5"= "#0f4926", "10" = "#0d2d1b", "50" = "#16261a" )
#grad_cols3 <- c("0"= "#1ee19b", "18.75" = "#2cbd89", "37.50" = "#0f9774", "62.50" = "#009657", "125" = "#226346", "250"= "#0f4926", "500" = "#0d2d1b", "2500" = "#16261a" ) #this is option if we want in µM instead of mM

grad_shapes <- c("0"= 4, "0.375" = 3, "0.75" = 5, "1.25" = 2, "2.5" = 0, "5"= 6, "10" = 10, "50" = 19)
#grad_shapes <- c("0"= 4, "18.75" = 3, "37.50" = 5, "62.50" = 2, "125" = 0, "250"= 6, "500" = 10, "2500" = 19) #this is option if we want in µM instead of mM

#tiff("Figures/Figure_1.tiff",width = 90,height = 30,res = 1000,units = "mm")
ggplot(h2o2.sum.reduced, aes(x = time_min, y = mean, color = H2O2_conc, shape = H2O2_conc))+
  geom_point(size = 0.8,stroke=0.3)+
  scale_shape_manual(values = grad_shapes,
                     name = bquote("H"[2]*"O"[2]~"mM"))+
  scale_colour_manual(values = grad_cols3,
                      name = bquote("H"[2]*"O"[2]~"mM"))+
  #scale_colour_brewer(palette = "PuBuGn") +
  theme_classic()+
  theme(
    #aspect.ratio = 0.7,
    legend.text.align = 0,
    text = element_text(size=7),
    legend.key.height = unit(0.1,'cm'))+
  guides(color=guide_legend(override.aes = list(size=1)))+
  xlim(0, 20)+
  ylab("Absorbance 590 nm")+
  xlab("Incubation time (min)")+
 facet_grid(~ MnP_conc, scales = "free")

```
## Homogenization test

Here, we tested the hypothesis that more intense sample homogenization would impact the assay by releasing internal peroxidases to the extract. We grew *Hypholoma fasiculare* on needles for about a month and tested the difference between three levels of extraction intensity. - Soaking in acetate buffer - Vortex shaking in acetate buffer (no visible tears in the mycelium) - Complete disruption of mycelium (UltraTurrax)

```{r wrangling_needle_data, echo = FALSE, message = FALSE, warning = FALSE}

# Load data
mbth_needle.dat <- read_csv("Data/Clean/clean_MBTH_needle_masceration.csv")
abts_needle.dat <- read_csv("Data/Clean/clean_ABTS_needle_masceration.csv")
ldopa_needle.dat <- read_csv("Data/Clean/clean_LDOPA_needle_masceration.csv")
dmp_needle.dat <- read_csv("Data/Clean/clean_DMP_needle_masceration.csv")

# Summarize mean values for technical replicates for each time point

mbth_needle.sum <- mbth_needle.dat |> 
  separate(sample,
           into = c("method", "replicate")) |>
  filter(method %in% c("shake", "soak", "turrax")) |> 
  group_by(time_min, method, treatment) |> 
  summarise(mean = mean(absorbance)) |> 
  mutate(substrate = "DMAB/MBTH")
mbth_needle.sum$method <- fct_relevel(mbth_needle.sum$method, "soak", "shake", "turrax")
mbth_needle.sum$treatment <- fct_relevel(mbth_needle.sum$treatment, "A_H2O2", "A", "B_H2O2", "B")


abts_needle.sum <- abts_needle.dat |> 
  separate(sample,
           into = c("method", "replicate")) |> 
  group_by(time_min, method, treatment) |> 
  summarise(mean = mean(absorbance))|> 
  mutate(substrate = "ABTS")
abts_needle.sum$method <- fct_relevel(abts_needle.sum$method, "soak", "shake", "turrax")
abts_needle.sum$treatment <- fct_relevel(abts_needle.sum$treatment, "A_H2O2", "A", "B_H2O2", "B")

ldopa_needle.sum <- ldopa_needle.dat |> 
  separate(sample,
           into = c("method", "replicate")) |> 
  group_by(time_min, method, treatment) |> 
  summarise(mean = mean(absorbance)) |> 
  mutate(substrate = "L-DOPA")
ldopa_needle.sum$method <- fct_relevel(ldopa_needle.sum$method, "soak", "shake", "turrax")
ldopa_needle.sum$treatment <- fct_relevel(ldopa_needle.sum$treatment, "A_H2O2", "A", "B_H2O2", "B")

dmp_needle.sum <- dmp_needle.dat |> 
  separate(sample,
           into = c("method", "replicate")) |> 
  group_by(time_min, method, treatment) |> 
  summarise(mean = mean(absorbance)) |> 
  mutate(substrate = "2,6-DMP")
dmp_needle.sum$method <- fct_relevel(dmp_needle.sum$method, "soak", "shake", "turrax")
dmp_needle.sum$treatment <- fct_relevel(dmp_needle.sum$treatment, "A_H2O2", "A", "B_H2O2", "B")

needle.sum.sup <- rbind(abts_needle.sum, ldopa_needle.sum, dmp_needle.sum)

needle.sum.sup <- needle.sum.sup |> 
  mutate(method = recode(method,
                         "soak" = "Soaking",
                         "shake" = "Shaking",
                         "turrax" = "Fine homogenization"))
mbth_needle.sum <- mbth_needle.sum |> 
  mutate(method = recode(method,
                         "soak" = "Soaking",
                         "shake" = "Shaking",
                         "turrax" = "Fine homogenization"))

```

**Figure 2** Absorbance plotted over time (15 min). Different extraction intensities and all four substrates. Points are mean values of 4 technical replicates from 3 biological replicates.

```{r needle_fig, echo = FALSE, message = FALSE, warning = FALSE}
needle.soak.mbth<-ggplot(mbth_needle.sum[which(mbth_needle.sum$method == "Soaking"),], aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment),shape = 16,size=0.8)+
  scale_colour_manual(values = custom_cols,
                      labels = custom_names,
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.7,
    text = element_text(size=8),
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2),direction = "horizontal"))+
  xlim(0,15)+
  ylim(0,1.5)+
  ylab("Absorbance")+
  xlab("")

needle.shak.mbth<-ggplot(mbth_needle.sum[which(mbth_needle.sum$method == "Shaking"),], aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment),shape = 16,size=0.8)+
  scale_colour_manual(values = custom_cols,
                      labels = custom_names,
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.7,
    legend.text.align = 0,
    text = element_text(size=8),
    #legend.position = "bottom",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2),direction = "horizontal"))+
  xlim(0,15)+
  ylim(0,1.5)+
  ylab("")+
  xlab("Incubation time (min)")

needle.homo.mbth<-ggplot(mbth_needle.sum[which(mbth_needle.sum$method == "Fine homogenization"),], aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment),shape = 16,size=0.8)+
  scale_colour_manual(values = custom_cols,
                      labels = custom_names,
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.7,
    text = element_text(size=8),
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2),direction = "horizontal"))+
  xlim(0,15)+
  ylim(0,1.5)+
  ylab("")+
  xlab("")

mn.indep.mbth<-ggplot(Mn_dependancy[which(Mn_dependancy$substrate == "MBTH/DMAB"),], aes(x=method,y=ratio,colour=substrate))+
  geom_jitter(position = position_dodge(0.6),size=1)+labs(x="Sample treatment",y="Proportion of \n Mn-independent  activity")+
  theme_classic()+
  scale_color_manual(values = c("#6567B8"))+
  theme(legend.position = "none",
        aspect.ratio = 0.7,
        legend.title = element_blank(),
        text = element_text(size = 8),
    plot.margin = unit(c(0,0,0,0),"cm"))+
  scale_x_discrete(labels=function(x) str_wrap(x,width = 5))+
 stat_summary_bin(data= Mn_dependancy[which(Mn_dependancy$substrate == "MBTH/DMAB"),] ,aes(x=method,y=ratio,colour=substrate,group=substrate),fun.data="mean_se",fun.args = list(mult=1), geom="errorbar",width=0.1, position = position_dodge(0.6),linewidth=0.3)+ #add the se of the mean 
  stat_summary_bin(data= Mn_dependancy[which(Mn_dependancy$substrate == "MBTH/DMAB"),] ,aes(x=method,y=ratio,colour=substrate,group=substrate),fun ="mean",geom = "point", position = position_dodge(0.6),size=2,shape=1) # add the mean points 

#combine figures
react<-(needle.soak.mbth+needle.shak.mbth+needle.homo.mbth)/guide_area()+plot_layout(guides = "collect")

layout.q<-"
AAAC
AAAC
AAAC
AAAC
BBB#
"
#tiff("Figures/Figure_2.tiff",width = 190,height = 60,res = 1000,units = "mm")
react+mn.indep.mbth+plot_layout(design = layout.q)+plot_annotation(tag_levels="a")
```


The more intense extraction method, the higher enzyme activity - this is not a big surprise. With the most intense extraction "turrax" there is some indication that H<sub>2</sub>O<sub>2</sub> was released in the solution. Absorbance decreased over time also with only solutions A or B in the reaction (no added H<sub>2</sub>O<sub>2</sub>).

More interestingly, the increase in absorbance in the "B + H<sub>2</sub>O<sub>2</sub>" treatment for homogenised samples suggests that some other type of peroxidase was released. I'm not sure whether the *Hypholoma* produces extracellular non-Mn dependent peroxidases but if it does, we should be able to observe an increase in absorbance in "B + H<sub>2</sub>O<sub>2</sub>" also in the shaken sample. I therefore assume that some internal peroxidases were released when we homogenized the mycelium. It is a bit difficult to interpret the significance of the non-Mn dependent peroxidases in relation to MnP in this assay since recovery of MnP also increased in that treatment.

We also did a non-replicated get an idea about how much enzyme could still be extracted after an initial round of shaking the plates. We therefore added more acetate buffer to the plates that had been shaken and tried the different levels of extraction intensity. It is clear that there is a lot of enzyme remaining after the first extraction round. This would mean that the extraction method should be standardized, otherwise semi-quantitative measurements would be difficult.

```{r resampling, echo = FALSE, message = FALSE, warning = FALSE}
#wrangle data

mbth_needle_resamp.sum <- mbth_needle.dat |> 
    separate(sample,
           into = c("method", "replicate")) |>
  filter(method %in% c("shakeshake", "shakesoak", "shaketurrax")) |>
  select(-replicate) |> 
  group_by(time_min, method, treatment) |> 
  summarise(mean = mean(absorbance)) |> 
  mutate(substrate = "DMAB/MBTH")
mbth_needle_resamp.sum$method <- fct_relevel(mbth_needle_resamp.sum$method, "shakesoak", "shakeshake", "shaketurrax")
mbth_needle_resamp.sum$treatment <- fct_relevel(mbth_needle_resamp.sum$treatment, "A_H2O2", "A", "B_H2O2", "B")

mbth_needle_resamp.sum <- mbth_needle_resamp.sum |> 
  mutate(method = recode(method,
                "shakesoak" = "Shaking + Soaking",
                "shakeshake" = "Shaking + Shaking",
                "shaketurrax" = "Shaking + fine homogenization"
                ))
```


**Supplemental Figure 1** Absorbance over time (MBTH/DMAB assay) of re-extracted plates. Absorbance values are slightly lower, but still similar to numbers obtained during first-round extractions.

```{r re_extract_needle_fig_sup,  echo = FALSE, message = FALSE, warning = FALSE}
needle.resoak.mbth<-ggplot(mbth_needle_resamp.sum[which(mbth_needle_resamp.sum$method == "Shaking + Soaking"),], aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment),shape = 16,size=0.8)+
  scale_colour_manual(values = custom_cols,
                      labels = custom_names,
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.7,
    text = element_text(size=8),
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2),direction = "horizontal"))+
  xlim(0,15)+
  ylim(0,1.5)+
  ylab("Absorbance")+
  xlab("")

needle.reshak.mbth<-ggplot(mbth_needle_resamp.sum[which(mbth_needle_resamp.sum$method == "Shaking + Shaking"),], aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment),shape = 16,size=0.8)+
  scale_colour_manual(values = custom_cols,
                      labels = custom_names,
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.7,
    legend.text.align = 0,
    text = element_text(size=8),
    #legend.position = "bottom",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2),direction = "horizontal"))+
  xlim(0,15)+
  ylim(0,1.5)+
  ylab("")+
  xlab("Incubation time (min)")

needle.rehomo.mbth<-ggplot(mbth_needle_resamp.sum[which(mbth_needle_resamp.sum$method == "Shaking + fine homogenization"),], aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment),shape = 16,size=0.8)+
  scale_colour_manual(values = custom_cols,
                      labels = custom_names,
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.7,
    text = element_text(size=8),
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2),direction = "horizontal"))+
  xlim(0,15)+
  ylim(0,1.5)+
  ylab("")+
  xlab("")

#tiff("Figures/Sup_Figure_1.tiff",width = 140,height = 60,res = 1000,units = "mm")
(needle.resoak.mbth+needle.reshak.mbth+needle.rehomo.mbth)/guide_area()+plot_layout(guides = "collect")+plot_annotation(tag_levels = "a")
```


**Supplemental Figure - all substrates needles test**

```{r needle_fig_sup, echo = FALSE, message = FALSE, warning = FALSE}
#tiff("Figures/Sup_Figure_2.tiff",width = 120,height = 180,res = 1000,units = "mm")
ggplot(needle.sum.sup, aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment),shape = 16,size=0.4)+
  scale_colour_manual(values = custom_cols,
                      labels = custom_names,
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.7,
    legend.text.align = 0,
    text = element_text(size=8),
    legend.position = "bottom")+
  guides(color=guide_legend(override.aes = list(size=2)))+
  xlim(0,15)+
  ylab("Absorbance")+
  xlab("Incubation time (min)")+
 facet_grid(substrate ~ method, scales = "free")
```


In order to be able to assess what impact the homogenization method has on the release of internal (Mn-independant) peroxidases and how this measurement varies between substrates we extracted the slopes from each reaction and compared the ratios of Mn-dependant to Mn-independant. Each slope is measured from the first 15 minutes of the reaction, except for in a few case where the range of time is adjusted to avoid stochastic points, i.e. slope taken over 1-15 minutes if 0-1 minutes were decreasing. 

```{r read_in_clean_data, echo = FALSE, message = FALSE, warning = FALSE}
#read in clean dataframe with ratios of Mn-dep to Mn-indep for each replicate for every substrate type, and method of homogenization. 
Mn_dependancy<-read_csv("Data/Clean/Mn_dependancy_ratios_clean.csv",col_types = list(col_factor(),col_factor(),col_factor(),col_number()))

Mn_dependancy <- Mn_dependancy |> 
  mutate(substrate = recode(substrate,
                "DMP" = "2,6-DMP"
                ))
Mn_dependancy$method <- fct_relevel(Mn_dependancy$method, "Soaking", "Shaking", "Fine homogenization")
```

**Sup Figure** Ratio between rate of activity for Mn-dependant and Mn-independant reactions. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#this is the figure when ABTS is included
#tiff("Figures/Sup_Figure_3.tiff",width = 170,height = 100,res = 1000,units = "mm")
ggplot(Mn_dependancy, aes(x=method,y=ratio,colour=substrate))+
  geom_jitter(position = position_dodge(0.6),size=2)+labs(x="Sample treatment",y="Proportion of Mn-independent activity")+
  theme_classic()+
  scale_color_manual(values = c("#6567B8","#96C5AB","#E7BB74","#EE896D"))+
  theme(legend.position = "right",legend.title = element_blank(),axis.text = element_text(size = 13))+
 stat_summary_bin(data= Mn_dependancy ,aes(x=method,y=ratio,colour=substrate,group=substrate),fun.data="mean_se",fun.args = list(mult=1), geom="errorbar",width=0.1, position = position_dodge(0.6),linewidth=0.5)+ #add the se of the mean 
  stat_summary_bin(data= Mn_dependancy ,aes(x=method,y=ratio,colour=substrate,group=substrate),fun ="mean",geom = "point", position = position_dodge(0.6),size=2,shape=1)+# add the mean points
  geom_hline(yintercept = 1,linetype="dashed")  


##substrates are colour coded according to the colour the substrate produces when oxidized
```


## Mn-dependency

We tested the assumption that the MnSO<sub>4</sub> and EDTA controls could be used to distinguish between Mn-dependent and independent oxidation of the four different substrates (ABTS, DMP, L-DOPA, MBTH/DMAB). All assay conditions were identical except for the dye-forming substrate. We ran incubations with either recombinant manganese peroxidase (MnP) from \*Phanerochaete\* or a versitile peroxidase from horseradish (HoP). In these assays, we used a lower H<sub>2</sub>O<sub>2</sub> concentration (2.5 mM) than we have used previously (5 mM in Kyaschenko protocol).

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Load data from 2023-12-05
Mn_depend.dat <- read_csv("Data/Clean/clean_MnP_HoP_comparison.csv")

# Summarize mean absorbance
Mn_depend.sum <- Mn_depend.dat |> 
  group_by(enzyme, time_min, treatment, substrate) |> 
  summarise(mean = mean(absorbance)) 
Mn_depend.sum$treatment <- fct_relevel(Mn_depend.sum$treatment, "A_H2O2", "A", "B_H2O2", "B")

# change enzyme abbreviation from HoP to HRP for figure, change MBTH/DMAB to DMAB/MBTH
Mn_depend.sum$enzyme[]<-sapply(Mn_depend.sum$enzyme,function(x) gsub("HoP","HRP",x))
Mn_depend.dat$enzyme[]<-sapply(Mn_depend.dat$enzyme,function(x) gsub("HoP","HRP",x))
Mn_depend.sum$substrate[]<-sapply(Mn_depend.sum$substrate,function(x) gsub("MBTH/DMAB","DMAB/MBTH",x))
Mn_depend.dat$substrate[]<-sapply(Mn_depend.dat$substrate,function(x) gsub("MBTH/DMAB","DMAB/MBTH",x))
Mn_depend.sum$substrate[]<-sapply(Mn_depend.sum$substrate,function(x) gsub("DMP","2,6-DMP",x))
Mn_depend.dat$substrate[]<-sapply(Mn_depend.dat$substrate,function(x) gsub("DMP","2,6-DMP",x))

```

**Supplemental figure 4** Absorbance plotted over time (30 min incubation) of all substrates and either MnP or HoP. Concentrations of MnP and HoP seem to be balanced for the MBTH/DMAB assay (similar absorbance). However, the other substrates seem to diverge in their specificity for the different enzymes. DMP and L-DOPA are easily oxidized by MnP but not by HoP while ABTS is rapidly oxidised by HoP but not by MnP.


```{r plotting_dependency_MnP_vs_HoP_new, echo = FALSE, message = FALSE, warning = FALSE}
react_plots<-ggplot(Mn_depend.sum, aes(x = time_min, y = mean))+
  geom_point(aes(colour = treatment), shape = 16,size=0.4)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 0.5,
    legend.text.align = 0,
    text=element_text(size=8),
    legend.key.size = unit(0.98,"line"),
    legend.position = "bottom",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  xlim(0, 15)+
  ylab("Absorbance")+
  xlab("Incubation time (min)")+
 facet_grid(substrate ~ enzyme, scales = "free")+
 geom_vline(xintercept = 5,linetype="dashed")
```

```{r read_in_clean_data, echo = FALSE, message = FALSE, warning = FALSE}
#read in data
activity_all_treatments<-read_csv("Data/clean/MnP_vs_HRP_extracted_slopes.csv")
```

```{r mbth_sup_4, echo = FALSE, message = FALSE, warning = FALSE}
inset.mbth.hrp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "HRP" & activity_all_treatments$substrate == "MBTH/DMAB"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    axis.title = element_text(size = 8),
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("")+
  xlab("Assay Control")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")

inset.mbth.mnp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "MnP" & activity_all_treatments$substrate == "MBTH/DMAB"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    axis.title = element_text(size = 8),
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("Abs/min")+
  xlab("Assay Control")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")

```

```{r abts_sup_4, echo = FALSE, message = FALSE, warning = FALSE}

inset.abts.hrp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "HRP" & activity_all_treatments$substrate == "ABTS"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("")+
  xlab("")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")

inset.abts.mnp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "MnP" & activity_all_treatments$substrate == "ABTS"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("Abs/min")+
  xlab("")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")

```

```{r dopa_sup_4, echo = FALSE, message = FALSE, warning = FALSE}
inset.dopa.hrp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "HRP" & activity_all_treatments$substrate == "L-DOPA"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("")+
  xlab("")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")

inset.dopa.mnp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "MnP" & activity_all_treatments$substrate == "L-DOPA"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("Abs/min")+
  xlab("")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")

```

```{r dmp_sup_4, echo = FALSE, message = FALSE, warning = FALSE}
inset.dmp.hrp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "HRP" & activity_all_treatments$substrate == "DMP"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("")+
  xlab("")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")


inset.dmp.mnp<-ggplot(activity_all_treatments[which(activity_all_treatments$treatment %in% c("A_H2O2","B_H2O2") & activity_all_treatments$enzyme == "MnP" & activity_all_treatments$substrate == "DMP"),], aes(x = treatment, y = slope))+
  geom_point(aes(colour = treatment), shape = 19,size=2)+
  scale_colour_manual(values = custom_cols, 
                      labels = custom_names, 
                      name = "Assay controls")+
  theme_classic()+
  theme(
    aspect.ratio = 1.2,
    text=element_text(size=8),
    axis.text.x = element_blank(),
    legend.position = "none",
    plot.margin = unit(c(0,0,0,0),"cm"))+
  guides(color=guide_legend(override.aes = list(size=2)))+
  ylab("Abs/min")+
  xlab("")+
  ylim(0, NA)+
  coord_cartesian(clip = "off")

```

```{r sup_4_all_panels, echo = FALSE, message = FALSE, warning = FALSE}
insets.patched<-inset.abts.mnp+inset.abts.hrp+inset.dmp.mnp+inset.dmp.hrp+inset.dopa.mnp+inset.dopa.hrp+inset.mbth.mnp+inset.mbth.hrp+plot_layout(nrow = 4)
patched<-react_plots|insets.patched
#tiff("Figures/Sup_Figure_4.tiff",width = 200,height = 120,res = 1000,units = "mm")
patched+plot_annotation(tag_levels = 'a')
```



Main conclusions: - Cannot distinguish Mn-dependent from Mn-independent oxidation with ABTS - Lack of linearity is a problem with the Mn-control. There seems to be some shift in equilibria between reagents/products in the reactions (see "blip" at \~10 min)
